---
title: "ChIP-seq data from ENCODE"
author: "Kaixuan Luo"
date: 2018-04-09
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## Download BAM files from ENCODE, then get 5' end counts in bigwig format
* bam files are in `/project/mstephens/data_smoothing/chipseq/bam/`
* bw files are in `/project/mstephens/data_smoothing/chipseq/bw/`
```{bash eval=FALSE}

mkdir /project/mstephens/data_smoothing/chipseq/log
cd /project/mstephens/data_smoothing/chipseq/log

mkdir /project/mstephens/data_smoothing/chipseq/bw

## CTCF HelaS3 Broad ChIP-seq
sbatch ~/projects/data_smoothing/data_smoothing_workflow/code/count_5end_coverage_bamToBigwig.sbatch /project/mstephens/data_smoothing/chipseq/bam/wgEncodeBroadHistoneHelas3CtcfStdAlnRep1.bam /project/mstephens/data_smoothing/chipseq/bw/wgEncodeBroadHistoneHelas3CtcfStdAlnRep1

sbatch ~/projects/data_smoothing/data_smoothing_workflow/code/count_5end_coverage_bamToBigwig.sbatch /project/mstephens/data_smoothing/chipseq/bam/wgEncodeBroadHistoneHelas3CtcfStdAlnRep2.bam /project/mstephens/data_smoothing/chipseq/bw/wgEncodeBroadHistoneHelas3CtcfStdAlnRep2

## CTCF HelaS3 UTA ChIP-seq
sbatch ~/projects/data_smoothing/data_smoothing_workflow/code/count_5end_coverage_bamToBigwig.sbatch /project/mstephens/data_smoothing/chipseq/bam/wgEncodeOpenChromChipHelas3CtcfAlnRep1.bam /project/mstephens/data_smoothing/chipseq/bw/wgEncodeOpenChromChipHelas3CtcfAlnRep1

sbatch ~/projects/data_smoothing/data_smoothing_workflow/code/count_5end_coverage_bamToBigwig.sbatch /project/mstephens/data_smoothing/chipseq/bam/wgEncodeOpenChromChipHelas3CtcfAlnRep2.bam /project/mstephens/data_smoothing/chipseq/bw/wgEncodeOpenChromChipHelas3CtcfAlnRep2

```


## Extract counts from bigwig files
```{bash}
cat /project/mstephens/data_smoothing/regions/example_CTCF_MACE.bed
```

* read counts are in `/project/mstephens/data_smoothing/chipseq/examples/`
```{bash eval=FALSE}

dir_examples=/project/mstephens/data_smoothing/chipseq/examples/
mkdir -p ${dir_examples}

dir_bw=/project/mstephens/data_smoothing/chipseq/bw

## CTCF HelaS3 Broad ChIP-seq
bwtool extract bed /project/mstephens/data_smoothing/regions/example_CTCF_MACE.bed ${dir_bw}/wgEncodeBroadHistoneHelas3CtcfStdAlnRep1_forward.bw ${dir_examples}/example_CTCF_MACE_wgEncodeBroadHistoneHelas3CtcfStdAlnRep1_forward_counts.txt

bwtool extract bed /project/mstephens/data_smoothing/regions/example_CTCF_MACE.bed ${dir_bw}/wgEncodeBroadHistoneHelas3CtcfStdAlnRep1_reverse.bw ${dir_examples}/example_CTCF_MACE_wgEncodeBroadHistoneHelas3CtcfStdAlnRep1_reverse_counts.txt

bwtool extract bed /project/mstephens/data_smoothing/regions/example_CTCF_MACE.bed ${dir_bw}/wgEncodeBroadHistoneHelas3CtcfStdAlnRep2_forward.bw ${dir_examples}/example_CTCF_MACE_wgEncodeBroadHistoneHelas3CtcfStdAlnRep2_forward_counts.txt

bwtool extract bed /project/mstephens/data_smoothing/regions/example_CTCF_MACE.bed ${dir_bw}/wgEncodeBroadHistoneHelas3CtcfStdAlnRep2_reverse.bw ${dir_examples}/example_CTCF_MACE_wgEncodeBroadHistoneHelas3CtcfStdAlnRep2_reverse_counts.txt

## CTCF HelaS3 UTA ChIP-seq
bwtool extract bed /project/mstephens/data_smoothing/regions/example_CTCF_MACE.bed ${dir_bw}/wgEncodeOpenChromChipHelas3CtcfAlnRep1_forward.bw ${dir_examples}/example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep1_forward_counts.txt

bwtool extract bed /project/mstephens/data_smoothing/regions/example_CTCF_MACE.bed ${dir_bw}/wgEncodeOpenChromChipHelas3CtcfAlnRep1_reverse.bw ${dir_examples}/example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep1_reverse_counts.txt

bwtool extract bed /project/mstephens/data_smoothing/regions/example_CTCF_MACE.bed ${dir_bw}/wgEncodeOpenChromChipHelas3CtcfAlnRep2_forward.bw ${dir_examples}/example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep2_forward_counts.txt

bwtool extract bed /project/mstephens/data_smoothing/regions/example_CTCF_MACE.bed ${dir_bw}/wgEncodeOpenChromChipHelas3CtcfAlnRep2_reverse.bw ${dir_examples}/example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep2_reverse_counts.txt

```

## read read counts
```{r}

extract_counts <- function(filename){
  bed_counts <- read.table(filename, header = F, stringsAsFactors = F)
  colnames(bed_counts) <- c("chr", "start", "end", "name", "width", "counts")
  
  counts <- strsplit(bed_counts$counts, split = ",")[[1]]
  counts[counts == "NA"] <- 0
  counts <- as.numeric(counts)

  return(counts.l = list(chr = bed_counts$chr, start = bed_counts$start, end = bed_counts$end, counts = counts))
}

dir_examples <- "/project/mstephens/data_smoothing/chipseq/examples/"

## Broad
counts_broad_rep1_forward.l <- extract_counts(file.path(dir_examples, "example_CTCF_MACE_wgEncodeBroadHistoneHelas3CtcfStdAlnRep1_forward_counts.txt"))

counts_broad_rep2_forward.l <- extract_counts(file.path(dir_examples, "example_CTCF_MACE_wgEncodeBroadHistoneHelas3CtcfStdAlnRep2_forward_counts.txt"))

counts_broad_rep1_reverse.l <- extract_counts(file.path(dir_examples, "example_CTCF_MACE_wgEncodeBroadHistoneHelas3CtcfStdAlnRep1_reverse_counts.txt"))

counts_broad_rep2_reverse.l <- extract_counts(file.path(dir_examples, "example_CTCF_MACE_wgEncodeBroadHistoneHelas3CtcfStdAlnRep2_reverse_counts.txt"))

## UTA
counts_UTA_rep1_forward.l <- extract_counts(file.path(dir_examples, "example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep1_forward_counts.txt"))

counts_UTA_rep2_forward.l <- extract_counts(file.path(dir_examples, "example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep2_forward_counts.txt"))

counts_UTA_rep1_reverse.l <- extract_counts(file.path(dir_examples, "example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep1_reverse_counts.txt"))

counts_UTA_rep2_reverse.l <- extract_counts(file.path(dir_examples, "example_CTCF_MACE_wgEncodeOpenChromChipHelas3CtcfAlnRep1_reverse_counts.txt"))
```

## plot CTCF Hela ChIP-seq read counts (Broad)
```{r, fig.height=10, fig.width=10}

title <- paste0(counts_broad_rep1_forward.l$chr, ": ", counts_broad_rep1_forward.l$start, "-",counts_broad_rep1_forward.l$end)

par(mfcol = c(4,1))
par(mar = c(2,2,2,1))
par(bty="n")
plot(counts_broad_rep1_forward.l$counts, main = title, type = "h", col = "blue", ylab = "rep1 forward", xlab = "", ylim = c(0,20))
plot(counts_broad_rep2_forward.l$counts, type = "h", col = "blue", ylab = "rep2 forward", xlab = "", ylim = c(0,20))
plot(counts_broad_rep1_reverse.l$counts, main = title, type = "h", col = "red", ylab = "rep1 reverse", xlab = "", ylim = c(0,20))
plot(counts_broad_rep2_reverse.l$counts, type = "h", col = "red", ylab = "rep2 reverse", xlab = "", ylim = c(0,20))

```


## plot CTCF Hela ChIP-seq read counts (UT-A)
```{r, fig.height=10, fig.width=10}

title <- paste0(counts_UTA_rep1_forward.l$chr, ": ", counts_UTA_rep1_forward.l$start, "-",counts_UTA_rep1_forward.l$end)

par(mfcol = c(4,1))
par(mar = c(2,2,2,1))
par(bty="n")
plot(counts_UTA_rep1_forward.l$counts, main = title, type = "h", col = "blue", ylab = "rep1 forward", xlab = "", ylim = c(0,20))
plot(counts_UTA_rep2_forward.l$counts, type = "h", col = "blue", ylab = "rep2 forward", xlab = "", ylim = c(0,20))
plot(counts_UTA_rep1_reverse.l$counts, main = title, type = "h", col = "red", ylab = "rep1 reverse", xlab = "", ylim = c(0,20))
plot(counts_UTA_rep2_reverse.l$counts, type = "h", col = "red", ylab = "rep2 reverse", xlab = "", ylim = c(0,20))

```
## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
